Sky Tomography for 3D Aerosol Distribution Recovery

### 2013-07-26 article.tex
bookmark

%
% Some new commands I use in this text
%
\newcommand{\OpSphere}{\bm{\mathcal{S}}}
\newcommand{\OpRot}{\bm{\mathcal{R}}}
\newcommand{\OpDistance}{\bm{\mathcal{D}}}
\newcommand{\OpCumsum}{\bm{\mathcal{C}}}
\newcommand{\OpInt}{\bm{\mathcal{I}}}
\newcommand{\OpCamera}{\bm{\mathcal{P}}}
\newcommand{\Mask}{\bm{M}}
\newcommand{\OpDiag}[1]{\mathrm{diag}\left\{#1\right\}}
\newcommand{\Grad}[1]{\bm{\triangledown_{#1}}}
\newcommand{\argmin}{\mathrm{arg}\min}
\newcommand{\curly}[1]{\left\{#1\right\}}
\newcommand{\roundy}[1]{\left(#1\right)}
\newcommand{\recty}[1]{\left[#1\right]}
\newcommand{\PartDeriv}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\mat}[1]{\bm{#1}}
\newcommand{\transpose}[1]{{#1}^\intercal}
\newcommand{\derivsym}[1]{\,d{#1}}
\newcommand{\yoavcomment}[1]{}
\renewcommand{\yoavcomment}[1]{#1} % Comment to remove images

@@

The atmosphere is discretized into a grid of volume elements (voxels).
A forward model takes as input the aerosol distribution over the grid
of voxels, and outputs images as if captured at various viewpoints.
In a \emph{single-scattering} model, any SR changes direction only
once on the way to the camera. This model is valid in atmospheres that
are not very dense: inside fog and clouds this model does not apply.
Under single-scattering approximation, radiative transfer has three
steps (\cref{fig:groundgrid}):
\begin{inparaenum}[\hfill\bfseries a\upshape)]
\item \label{itm:first} Attenuation of a SR propagating from the top
  of the atmosphere (TOA) to voxel $k$.
\item \label{itm:second} Light scattering at voxel $k$, towards a
  camera.
\item \label{itm:third} Light attenuation on the LOS from voxel $k$ to
  the camera.
\end{inparaenum}
Steps \ref{itm:first} and \ref{itm:third} involve the optical depth
along the light path to and from voxel $k$, analyzed in
\cref{sec:optical-depth}.  Step \ref{itm:second} involves the
scattering coefficient at voxel $k$, described in
\cref{sec:scattering}.

@@

A voxel's vertical geometric thickness is $\Delta z$.  The Sun zenith
angle is $\Phi^{\rm SR}$. Denote a SR line segment between the TOA and
the center of voxel $k$ by $[{\rm SR},k]$. This SR intersects voxel
$q$. The geometric length of this intersection line segment is $l_{\rm
  SR}(q)$ which depends on $\Delta z$, $\Phi^{\rm SR}$ and $q$.
Following \cref{eq:tau,eq:betair}, the optical depth between the TOA
and voxel $k$ is
\begin{equation}
  \tau_{\rm SR}(k)= \sum_{q \in[{\rm SR},k]} \beta(q) l_{\rm SR}(q) \;\;.
  \label{eq:lSRn}
\end{equation}
Based on \cref{eq:betaer,eq:betair,eq:lSRn},
\begin{equation}
  \tau_{\rm SR}(k)=
  \tau_{\rm SR}^{\rm air}(k) +  \tau_{\rm SR}^{\rm aerosol}(k)
  \;\;,
  \label{eq:lSRaa}
\end{equation}
where
\begin{equation}
  \tau_{\rm SR}^{\rm air}(k)=
  \sum_{q \in[{\rm SR},k]}
  \beta^{\rm air}(q)  l_{\rm SR}(q)
  \label{eq:lSRair}
\end{equation}
and
\begin{equation}
  \tau_{\rm SR}^{\rm aerosol}(k)
  \sum_{q \in[{\rm SR},k]}
  \sigma^{\rm aerosol}(q) n^{\rm aerosol}(q) l_{\rm SR}(q)
  \label{eq:lSRaeros}
\end{equation}

The number of voxels is $N_{\rm voxels}$. All their associated
variables are concatenated into columns stack vectors, each $N_{\rm
  voxels}$ long. Aerosol densities and cross sections in all voxels
are stored in the respective vectors ${\bm n}$ and
$\vect{\sigma}$. Define a $N_{\rm voxels}\times N_{\rm voxels}$ sparse
matrix ${\bm D}^{\mathrm{Sun \rightarrow voxel}}$, whose element
$(k,q)$ is
\begin{equation}
  D^{\mathrm{Sun \rightarrow voxel}}(k,q) =
  \left\{
    \begin{array}{ll}
      l_{\rm SR}(q) & \mbox{ if $q \in[{\rm SR},k]$} \\
      0  & \mbox{ otherwise}
      \label{eq:Dsv}
    \end{array}
  \right.
  .
  \hspace{-0.05cm}
\end{equation}
\cref{eq:lSRaeros,eq:Dsv} lead to a vector of SR optical depths due to
aerosols (See \cref{fig:projection})a:
\begin{equation}
  \vect{\tau}_{\rm SR}^{\rm aerosol}=
  {\bm D}^{\mathrm{Sun \rightarrow voxel}}
  (\vect{\sigma} \odot {\bm n})
  \label{eq:tauSRvec1}
\end{equation}
where the operator $\odot$ denotes the Hadamard (element-wise)
product. Similarly, \cref{eq:lSRair} leads to a vector of SR optical
depths due to air, $\vect{\tau}_{\rm SR}^{\rm air}$.

Analogously, for camera $c$, denote by $[{\rm LOS}_c,k]$ a LOS bounded
between the camera and the center of voxel $k$. The LOS zenith angle
is $\Phi^{\rm LOS}_c$.  This LOS intersects voxel $q$. The geometric
length of this intersection line segment is $l_{\rm LOS_c}(q)$which
depends on $\Delta z$, $\Phi^{{\rm LOS}_c}$ and $q$.  Analogously to
\cref{eq:lSRaa,eq:lSRair,eq:lSRaeros}, the optical depth along a LOS
to $k$ is
\begin{equation}
  \tau_{{\rm LOS}_c}(k)=
  \tau_{{\rm LOS}_c}^{\rm air}(k) +  \tau_{{\rm LOS}_c}^{\rm aerosol}(k)
  \;\;,
  \label{eq:lLOSaerot}
\end{equation}
where
\begin{equation}
  \tau_{{\rm LOS}_c}^{\rm air}(k)=
  \sum_{q \in[{\rm LOS}_c,k]}
  \beta^{\rm air}(q)  l_{\rm LOS_c}(q)
  \label{eq:lLOSair}
\end{equation}
and
\begin{align}
  \label{eq:lLOSaeros}
  \tau_{{\rm LOS}_c}^{\rm aerosol}(k) = \sum_{q \in[{\rm LOS}_c,k]}
  \sigma^{\rm aerosol}(q) n^{\rm aerosol}(q) l_{\rm LOS_c}(q).
\end{align}
\cref{eq:lLOSaeros} leads to a vector of LOS optical depths due to
aerosols:
\begin{equation}
  \vect{\tau}_{{\rm LOS}_c}^{\rm aerosol}=
  {\bm D}^{\mathrm{voxel \rightarrow cam}}_c
  (\vect{\sigma} \odot {\bm n})
  \label{eq:tauLOSvec1}
\end{equation}
Here ${\bm D}^{\mathrm{voxel \rightarrow cam}}_c$ is a $N_{\rm
  voxels}\times N_{\rm voxels}$ sparse matrix whose element $(k,q)$ is
\begin{equation}
  D^{\mathrm{voxel \rightarrow cam}}_c(k,q) =
  \left\{
    \begin{array}{ll}
      l_{\rm LOS_c}(q) & \mbox{ if $q \in[{\rm LOS}_c,k]$} \\
      0  & \mbox{ otherwise}
      \label{eq:Dvc}
    \end{array}
  \right.
  .
  \hspace{-0.05cm}
\end{equation}
Similarly, \cref{eq:lSRair} leads to a vector of LOS optical depths
due to air, $\vect{\tau}_{{\rm LOS}_c}^{\rm air}$.

The fixed sparse matrices ${\bm D}^{\mathrm{Sun \rightarrow voxel}}$
and ${\bm D}^{\mathrm{voxel \rightarrow cam}}_c$ are compounded to a
single matrix
\begin{equation}
  {\bm D}_c=
  {\bm D}^{\mathrm{Sun \rightarrow voxel}}+
  {\bm D}^{\mathrm{voxel \rightarrow cam}}_c
  \;\;.
  \label{eq:Dtotal}
\end{equation}
Similarly, the molecular optical depths along all trajectories between
the sun, each voxel and camera $c$ are compounded to a vector
\begin{equation}
  \vect{\tau}_c^{\rm air}=
  \vect{\tau}_{\rm SR}^{\rm air}+
  \vect{\tau}_{{\rm LOS}_c}^{\rm air}
  \;\;.
  \label{eq:tauairtotal}
\end{equation}
Using \cref{eq:rayleighbeta} for $\beta^{\rm air}$ in all voxels, the
vector $\vect{\tau}_c^{\rm air}$ is fixed. Thus $\vect{\tau}_c^{\rm
  air}$ can be calculated once, per sun zenith angle.

Using
\cref{eq:lSRaa,eq:tauSRvec1,eq:lLOSair,eq:tauLOSvec1,eq:Dtotal,eq:tauairtotal},
the total optical depths corresponding to LOSs (of camera $c$) and SRs
that cross all voxels comprise the vector
\begin{equation}
  \vect{\tau}_c= \vect{\tau}_c^{\rm air}
  + {\bm D}_c (\vect{\sigma} \odot {\bm n})
  \;\;.
  \label{eq:tautotal}
\end{equation}
This vector depends on the variables $\vect{\sigma}$ and ${\bm n}$
(See~\cref{fig:projection}a).

@@

Per camera $c$ and voxel $k$, the lines $[{\rm SR},k]$ and $[{\rm
  LOS}_c,k]$ intersect at a fixed angle $\Phi^{\rm scatter}_c(k)$,
which is pre-calculated.  Column-stacking all voxels yields the vector
representation of all scattering angles in the domain,
$\vect{\Phi}^{\rm scatter}_c$ per camera $c$. Similarly, the
single-scattering albdeos $\varpi^{\rm aerosol}$ in all voxels are
stacked to vector $\vect{\varpi}$.  Using
\cref{eq:alphabasic,eq:rayleighP,eq:betaer,eq:betair}, the angular
scattering coefficients across the domain are expressed in vector form
by
\begin{align}
  \tilde{\vect{\alpha}}^{\rm air}_c &=
  \vect{\beta}^{\rm air} \odot P^{\rm air}(\vect{\Phi}^{\rm scatter}_c) \nonumber \\
  \tilde{\vect{\alpha}}^{\rm aerosol}_c &= \vect{\varpi}\odot
  \vect{\sigma} \odot {\bm n} \odot P^{\rm aerosol}_{\bm
    g}(\vect{\Phi}^{\rm scatter}_c)
  \label{eq:alpha_matrix}
\end{align}
The vector ${\bm g}$ column-stacks the anisotropy parameter of the
phase function, in all voxels. We use this notation, to indicate a
parametric phase function as in \cref{eq:aerosol_scatter}.

>>>>

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

%
% Some new commands I use in this text
%
\newcommand{\OpSphere}{\mathbfcal{S}}
\newcommand{\OpRot}{\mathbfcal{R}}
\newcommand{\OpDistance}{\mathbfcal{D}}
\newcommand{\OpCumsum}{\mathbfcal{C}}
\newcommand{\OpInt}{\mathbfcal{I}}
\newcommand{\OpCamera}{\mathbfcal{P}}
\newcommand{\Mask}{\bm{M}}
\newcommand{\OpDiag}[1]{\mathrm{diag}\left\{#1\right\}}
\newcommand{\Grad}[1]{\bm{\triangledown_{#1}}}
\newcommand{\argmin}{\mathrm{arg}\min}
\newcommand{\curly}[1]{\left\{#1\right\}}
\newcommand{\roundy}[1]{\left(#1\right)}
\newcommand{\recty}[1]{\left[#1\right]}
\newcommand{\PartDeriv}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\mat}[1]{\bm{#1}}
\newcommand{\transpose}[1]{{#1}^\intercal}
\newcommand{\derivsym}[1]{\,d{#1}}
\newcommand{\yoavcomment}[1]{}
\renewcommand{\yoavcomment}[1]{#1} % Comment to remove images

@@

The atmosphere is discretized into a grid of $N_{\rm voxels}$ volume
elements (voxels).  A forward model takes as input the aerosol
distribution over the grid of voxels, and outputs images as if
captured at various viewpoints.  In a \emph{single-scattering} model,
any SR changes direction only once on the way to the camera. This
model is valid in atmospheres that are not very dense: inside fog and
clouds this model does not apply.  Under single-scattering
approximation, radiative transfer has three steps
(\cref{fig:groundgrid}):
\begin{inparaenum}[\hfill\bfseries a\upshape)]
\item \label{itm:first} Attenuation of a SR propagating from the top
  of the atmosphere (TOA) to voxel $k$.
\item \label{itm:second} Light scattering at voxel $k$, towards a
  camera.
\item \label{itm:third} Light attenuation on the LOS from voxel $k$ to
  the camera.
\end{inparaenum}
Steps \ref{itm:first} and \ref{itm:third} involve the optical depth
along the light path to and from voxel $k$, analyzed in
\cref{sec:optical-depth}.  Step \ref{itm:second} involves the
scattering coefficient at voxel $k$, described in
\cref{sec:scattering}.

@@

A voxel's vertical geometric thickness is $\Delta z$.  The Sun zenith
angle is $\Phi^{\rm SR}$. Denote a SR line segment between the TOA and
the center of voxel $k$ by $[{\rm SR},k]$ (See
\cref{fig:groundgrid}). This SR intersects voxel $q$. The geometric
length of this intersection line segment is $l_{\rm SR}(q)$ which
depends on $\Delta z$, $\Phi^{\rm SR}$ and $q$.  We now define a
$N_{\rm voxels}\times N_{\rm voxels}$ matrix $\OpDistance^{\mathrm{Sun
    \rightarrow voxel}}$, whose element $(k,q)$ represents $l_{\rm
  SR}(q)$:
\begin{equation}
  \OpDistance^{\mathrm{Sun \rightarrow voxel}}(k,q) =
  \left\{
    \begin{array}{ll}
      l_{\rm SR}(q) & \mbox{ if $q \in[{\rm SR},k]$} \\
      0  & \mbox{ otherwise}
      \label{eq:Dsv}
    \end{array}
  \right.
  \hspace{-0.05cm}
\end{equation}
Note that matrix $\OpDistance^{\mathrm{Sun \rightarrow voxel}}$ is
sparse.  Following \cref{eq:tau}, the optical depth between the TOA
and voxel $k$ is
\begin{equation}
  \tau_{\rm SR}(k)= \sum_{q \in[{\rm SR},k]} \beta(q) l_{\rm SR}(q) \;\;.
  \label{eq:lSRn}
\end{equation}
We can write \cref{eq:lSRn} using matrix notation
\begin{equation}
  \vect{\tau}_{\rm SR}=\OpDistance^{\mathrm{Sun \rightarrow voxel}} \vect{\beta}
  \label{eq:tau_matrix}
\end{equation}
Here $\vect{\tau}_{\rm SR}$ and $\vect{\beta}$ are column stack vector
representation of $\tau_{\rm SR}(k)$ and $\beta(q)$, respectively.

Analogously, for camera $c$, denote by $[{\rm LOS}_c,k]$ a LOS bounded
between the camera and the center of voxel $k$ (see
\cref{fig:groundgrid}).  The LOS zenith angle is $\Phi^{\rm LOS}_c$.
This LOS intersects voxel $q$. The geometric length of this
intersection line segment is $l_{\rm LOS_c}(q)$which depends on
$\Delta z$, $\Phi^{{\rm LOS}_c}$ and $q$.  Analogously to
\cref{eq:Dsv}, we define a $N_{\rm voxels}\times N_{\rm voxels}$
sparse matrix whose element $(k,q)$ is
\begin{equation}
  \OpDistance^{\mathrm{voxel \rightarrow cam}}_c(k,q) =
  \left\{
    \begin{array}{ll}
      l_{{\rm LOS}_c}(q) & \mbox{ if $q \in[{\rm LOS}_c,k]$} \\
      0  & \mbox{ otherwise}
      \label{eq:Dvc}
    \end{array}
  \right.
  .
  \hspace{-0.05cm}
\end{equation}
The optical depth along a LOS to $k$ is
\begin{equation}
  \vect{\tau}_{{\rm LOS}_c} = \OpDistance^{\mathrm{voxel \rightarrow
      cam}} \vect{\beta}
  \label{eq:lLOSn}
\end{equation}
Here $\vect{\tau}_{\rm SR}$ is a column stack vector representation of
$\tau_{{\rm LOS}_c}(k)$.  The SR and LOS paths of each voxel $k$ can
be joined to a single path from the TOA to camera $c$ giving
\begin{equation}
  {\bm D}_c=
  {\bm D}^{\mathrm{Sun \rightarrow voxel}}+
  {\bm D}^{\mathrm{voxel \rightarrow cam}}_c
  \quad\text{ and }\quad
  \vect{\tau}_c=
  \vect{\tau}_{\rm SR}+
  \vect{\tau}_{{\rm LOS}_c}
  \;\;.
  \label{eq:D_tau_total}
\end{equation}

We discern between air molecules and other aerosols and therefore
write $ \vect{\tau}_c= \vect{\tau}_c^{\rm air} + \vect{\tau}_c^{\rm
  aerosol}$.  Using \cref{eq:rayleighbeta} for $\beta^{\rm air}$ in
all voxels, the vector $\vect{\tau}_c^{\rm air}$ is fixed. Thus
$\vect{\tau}_c^{\rm air}$ can be calculated once, per sun zenith
angle. Given \cref{eq:extinctc}, the total optical depths
corresponding to LOSs (of camera $c$) and SRs that cross all voxels
comprise the vector
\begin{equation}
  \vect{\tau}_c= \vect{\tau}_c^{\rm air}
  + {\bm D}_c (\vect{\sigma} \odot {\bm n})
  \;\;.
  \label{eq:tautotal}
\end{equation}
where the operator $\odot$ denotes the Hadamard (element-wise)
product. This vector depends on the variables $\vect{\sigma}$ and
${\bm n}$, the aerosol extinction and density at each voxel
respectively.

@@

Per camera $c$ and voxel $k$, the lines $[{\rm SR},k]$ and $[{\rm
  LOS}_c,k]$ intersect at a fixed angle $\Phi^{\rm scatter}_c(k)$.
Column-stacking all voxels yields the vector representation of all
scattering angles in the domain, $\vect{\Phi}^{\rm scatter}_c$ per
camera $c$. Similarly, the single-scattering albdeos $\varpi^{\rm
  aerosol}$ in all voxels are stacked to vector $\vect{\varpi}$.
Using \cref{eq:alphabasic,eq:rayleighP,eq:betaer,eq:betair}, the
angular scattering coefficients across the domain are expressed in
vector form by
\begin{align}
  \tilde{\vect{\alpha}}^{\rm air}_c &=
  \vect{\beta}^{\rm air} \odot P^{\rm air}(\vect{\Phi}^{\rm scatter}_c) \nonumber \\
  \tilde{\vect{\alpha}}^{\rm aerosol}_c &= \vect{\varpi}\odot
  \vect{\sigma} \odot {\bm n} \odot P^{\rm aerosol}_{\bm
    g}(\vect{\Phi}^{\rm scatter}_c)
  \label{eq:alpha_matrix}
\end{align}
The vector ${\bm g}$ column-stacks the anisotropy parameter of the
phase function, in all voxels. We use this notation, to indicate a
parametric phase function as in \cref{eq:aerosol_scatter}.
